---
title: "Handson for `tidyverse`"
author: Kazuharu Yanagimoto
date: '`r format(Sys.time(), "%B %d, %Y")`'
execute:
    warning: false
    message: false
    cache: false
format:
    html:
        self-contained: true
---

In this exercise, we are going to see a relationship between **fertility rate** and
**male's unpaid work**.

Both data is available in the World Bank Data:

- [Fertility rate](https://data.worldbank.org/indicator/SP.DYN.TFRT.IN)
- [Proportion of time spent on unpaid domestic and care work, male](https://data.worldbank.org/indicator/SG.TIM.UWRK.MA)

You can load them by the `WDI` package.

```{r}
library(tidyverse)
library(WDI)

fertility <- WDI(indicator = "SP.DYN.TFRT.IN", start = 2010, end = 2019) |>
  rename(fertility = "SP.DYN.TFRT.IN")
unpaid_male <- WDI(indicator = "SG.TIM.UWRK.MA", start = 2010, end = 2019) |>
  rename(hour_unpaid = "SG.TIM.UWRK.MA")
```

::: {.callout-warning}

## Quarto Notebook Cache

If you like to solve exercises by compiling the notebook,
set `cache: true` in the top yaml.

:::


## `dplyr` & `tidyr`

::: {.callout-tip}

## Q1 `glimpse`

Look through the data. Check

- What type of values are in the `country` column
- How many `NA` values are in the data

:::

```{r}
# Your Codes Here
```



::: {.callout-tip}

## Q2 `filter`

Let's focus only on OECD countries. 
Using the following vector `oecd`, choose only OECD-member countries.

```{r}
#remotes::install_github("caldwellst/whotilities")
oecd <- whoville::oecd_member_states()
```

:::

```{r}
# Your Codes Here
```

::: {.callout-tip}

## Q3 `group_by` & `summarize`

Now, we want to have a data point for each country.
Considering the data, compute

- Mean of fertility rate by country
- Latest data point in `unpaid_male` of each country (use `top_n()`)

I know the second one is tough. See the answer if you want.
:::

```{r}
# Your Codes Here
```

```{r}
# Your Codes Here
```


::: {.callout-tip}

## Q4 `left_join()`

Merge the two data frames you got in the last question.

:::

```{r}
# Your Codes Here
```

## `ggplot2`

::: {.callout-tip}

## Q5 `geom_point()`

Plot a scatter plot of male hours of unpaid work and fertility rate.
To make it better, try

- Fit a linear line
- Change the coordinates
- Pick another theme

:::


```{r}
# Your Codes Here
```


::: {.callout-tip}

## Q6 (Extra) `ggrepel::geom_text_repel()`

Put country labels on data points.
It is known that `ggrepel::geom_text_repel()` works better than the original `geom_text()`.

:::


```{r}
# Your Codes Here
```


